\textbf{Predicting virus-receptor mutant binding by molecular dynamics simulation}

### 2013-07-16 paper.tex
Small change in discussion.

\subsection*{Comparative analysis of the GP1/hTfR1 interface}
Considering the involvement of extended hydrogen-bonding networks in the GP1/hTfR1 interface, it was not clear that individual alanine mutations, even those that should destroy such networks, would significantly change the strength of interaction. One major advantage of first principles simulations is the ability to test mutations other than alanine without additional underlying assumptions in the energy function. Seen in Table~\ref{tab:summary_mutants}), we made additional mutations based on biochemical intuition or available experimental data to chemically diverse amino acids including tryptophan, lysine, aspartate, and threonine. Several mutations caused significant relative affinity changes. In addition, to detect synergistic effects, we tested several double mutants where both mutations appeared to cause similar changes in binding. Then, we compared the size of those differences to single mutants (Figure~\ref{fig:sensitivity} and ~\ref{fig:mutant_comparison}).

Although Tyr211Ala appears to have a large impact on binding affinity, no single mutant can provide enough evidence to understand the biochemical difference in binding mechanism. Since alanine is both smaller than tyrosine and also incapable of participating in hydrogen-bond interactions, we tested further mutations to identify the critical biochemical difference responsible for change in binding affinity. In particular, we substituted smaller side chains that, like tyrosine, were capable of hydrogen bonding. We chose Tyr211Asp and Tyr211Thr, two mutations that have been discussed in the context of selection pressure on hosts in rodent populations \citep{Rad2008,Rad20111,Rad20112}. Both mutations proved capable of causing a significant change in binding affinity in our simulations, but the change appeared to be increased affinity. (Figure~\ref{fig:sensitivity} and ~\ref{fig:mutant_comparison}, and Table~\ref{tab:MAF_p}).

We also simulated several point mutations at Asn348 in the hTfR1. As discussed above, the alanine mutation at this site showed no significant difference in maximum applied force or AUC from WT (Tables~\ref{tab:MAF_p} and ~\ref{tab:AUC_p}). In addition, the Asn348Lys mutation showed no significant difference from WT, and neither did Asn348Trp. (For both of these mutations, however, mean maximum applied force and mean AUC was lower than for WT. See Table~\ref{tab:summary_statistics}.)  On the other hand, there was a detectable difference between Asn348Ala and Asn348Lys (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), with Asn348Lys being a weaker binder. Moreover, Asn348Trp showed nearly identical results to Asn348Lys. The mutations to large amino acids (Asn348Trp and Asn348Lys) produced nearly identical affinity changes, whereas the mutations to amino acids not capable of hydrogen bonding (Asn348Ala and Asn348Trp) produced significantly different affinity changes (Table~\ref{tab:MAF_pairwise}). To check the consistency of our results, we hypothesized that the combination of Tyr211Ala and Asn348Trp, being chemically disconnected in two different hydrogen-bonding networks, would lead to a synergistic loss-of-binding. As expected, the double mutant was the weakest binding mutant tested ($ p < 10^{-6} $) (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}) in this study. Further, according to maximum applied force (but not AUC), the combination of Tyr211Ala and Asn348Trp also showed significantly weaker binding than Tyr211Ala by itself (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}). We suspect that the effect of Asn348Trp alone is near the limit of detection using our method. A larger number of replicates would possibly have resolved affinity differences between Asn348Trp and WT or other mutants more consistently.

@@

By extending the SMD technique to investigate a protein--protein interaction we have, in many ways, changed the scope of its application. Traditionally this technique is either applied to compute equilibrium quantities via a non-equilibrium approximation \citep{Park2003,Park2004,Giorgino2011}, used to estimate protein stability through unfolding \citep{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \citep{Dixit2001}. Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \citep{Cuendet2011,Giorgino2011}. And still others aimed at more esoteric problems, such as applying driving forces to untie polypeptide knots in toy systems \citep{Sulkowska2010}. Here, we have shown that SMD can provide insight into the \textit{relative} strength of protein--protein interactions. In contrast to more traditional absolute affinity calculations, relative binding is much more useful for experimental application. Further, our method complements whole cell infectivity assays, which demonstrate phenotypic effects but do not provide biochemical mechanisms. Via SMD, one can separate mutations whose likely effect is altered binding affinity from mutations that may affect other factors such as protein expression, protein folding, or interactions with other proteins not considered in the simulation. Thus, SMD may open avenues for subsequent experimental work.

Our findings may rationalize several effects observed in both infectivity data and rodent populations \citep{Rad2008,Rad20111}. First, we found that some substitutions at positions 211 and 348 did affect the strength of receptor binding. However, the computational data suggest that the reason and nature of the effects at these two sites are very different. Congruent with entry data \citep{Rad2008,Rad20111}, mutations at Tyr211 often have a dramatic effect on receptor binding. Second, no individual point mutation at Asn348 showed a significantly different affinity from the WT complex. However, when Tyr211Ala was paired with large (Trp) and positively charged (Lys) substitutions at position 348, the result was a larger than expected synergistic difference. That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually. Third, the GP1 mutation vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}, yet it was indistinguishable from the WT complex in our simulations. Although Tyr211Ala was the most disruptive single mutant we tested, vArg111Ala in the GP1 was able to restore mean maximum applied force to WT levels (Table~\ref{tab:summary_statistics}), and to levels significantly higher than observed for Tyr211Ala alone.

>>>>

\subsection*{Comparative analysis of the GP1/hTfR1 interface}
Considering the involvement of extended hydrogen-bonding networks in the GP1/hTfR1 interface, it was not clear that individual alanine mutations, even those that should destroy such networks, would significantly change the strength of interaction. One major advantage of first principles simulations is the ability to test mutations other than alanine without additional underlying assumptions in the energy function. Seen in Table~\ref{tab:summary_mutants}, we made additional mutations based on biochemical intuition or available experimental data to chemically diverse amino acids including tryptophan, lysine, aspartate, and threonine. Several mutations caused significant relative affinity changes. In addition, to detect synergistic effects, we tested several double mutants where both mutations appeared to cause similar changes in binding. Then, we compared the size of those differences to single mutants (Figure~\ref{fig:sensitivity} and ~\ref{fig:mutant_comparison}).

Although Tyr211Ala appears to have a large impact on binding affinity, no single mutant can provide enough evidence to understand the biochemical difference in binding mechanism. Since alanine is both smaller than tyrosine and also incapable of participating in hydrogen-bond interactions, we tested further mutations to identify the critical biochemical difference responsible for change in binding affinity. In particular, we substituted smaller side chains that, like tyrosine, were capable of hydrogen bonding. We chose Tyr211Asp and Tyr211Thr, two mutations that have been discussed in the context of selection pressure on hosts in rodent populations \citep{Rad2008,Rad20111,Rad20112}. Both mutations proved capable of causing a significant change in binding affinity in our simulations, but the change appeared to be increased affinity (Figure~\ref{fig:sensitivity} and ~\ref{fig:mutant_comparison}, and Table~\ref{tab:MAF_p}).

We also simulated several point mutations at Asn348 in the hTfR1. As discussed above, the alanine mutation at this site showed no significant difference in maximum applied force or AUC from WT (Tables~\ref{tab:MAF_p} and ~\ref{tab:AUC_p}). In addition, neither the Asn348Lys nor the Asn348Trp mutation showed a significant difference from WT. For both of these mutations, however, mean maximum applied force and mean AUC was lower than for WT (See Table~\ref{tab:summary_statistics}).  On the other hand, there was a detectable difference between Asn348Ala and Asn348Lys (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), with Asn348Lys being a weaker binder. Moreover, Asn348Trp showed nearly identical results to Asn348Lys. The mutations to large amino acids (Asn348Trp and Asn348Lys) produced nearly identical affinity changes, whereas the mutations to amino acids not capable of hydrogen bonding (Asn348Ala and Asn348Trp) produced significantly different affinity changes (Table~\ref{tab:MAF_pairwise}). To check the consistency of our results, we hypothesized that the combination of Tyr211Ala and Asn348Trp, being chemically disconnected in two different hydrogen-bonding networks, would lead to a synergistic loss-of-binding. As expected, the double mutant was the weakest binding mutant tested ($ p < 10^{-6} $) (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}) in this study. Further, according to maximum applied force (but not AUC), the combination of Tyr211Ala and Asn348Trp also showed significantly weaker binding than Tyr211Ala by itself (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}). We suspect that the effect of Asn348Trp alone is near the limit of detection using our method. A larger number of replicates would possibly have resolved affinity differences between Asn348Trp and WT or other mutants more consistently.

@@

By extending the SMD technique to investigate a protein--protein interaction we have, in many ways, changed the scope of its application. Traditionally this technique has been either applied to compute equilibrium quantities via a non-equilibrium approximation \citep{Park2003,Park2004,Giorgino2011}, used to estimate protein stability through unfolding \citep{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \citep{Dixit2001}. Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \citep{Cuendet2011,Giorgino2011}. And still others aimed at more esoteric problems, such as applying driving forces to untie polypeptide knots in toy systems \citep{Sulkowska2010}. Here, we have shown that SMD can provide insight into the \textit{relative} strength of protein--protein interactions. In contrast to more traditional absolute affinity calculations, relative binding is much more useful for experimental application. Further, our method complements whole cell infectivity assays, which demonstrate phenotypic effects but do not provide biochemical mechanisms. Via SMD, one can separate mutations whose likely effect is altered binding affinity from mutations that may affect other factors such as protein expression, protein folding, or interactions with other proteins not considered in the simulation. Thus, SMD may open avenues for subsequent experimental work.

Our findings may rationalize several effects observed in both infectivity data and rodent populations \citep{Rad2008,Rad20111}. First, we found that some substitutions at positions 211 and 348 did affect the strength of receptor binding. However, the computational data suggest that the reason and nature of the effects at these two sites are very different. At position 211, mutations to non-polar residues causes a large change in binding. This is congruent with what is known from viral entry data \citep{Rad2008,Rad20111}. By contrast, mutations at position 348 need only be small to maintain WT binding. The ability to hydrogen bond appears to be insignificant. This can be inferred from the fact that Tyr211Ala paired with large (Trp) and positively charged (Lys) substitutions at position 348 results in a larger than expected synergistic difference. That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually. Third, the GP1 mutation vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}, yet it was indistinguishable from the WT complex in our simulations. Although Tyr211Ala was the most disruptive single mutant we tested, vArg111Ala in the GP1 was able to restore mean maximum applied force to WT levels (Table~\ref{tab:summary_statistics}), and to levels significantly higher than observed for Tyr211Ala alone.
